Komols Complexity, Coresets, and Sketches in Machine Learning

### 2019-01-07 komlos.tex
edits

\begin{definition}
Let $\F$ be a family of functions $f:\X\rightarrow\R$ and $\sigma \in \{-1,1\}^m$ uniformly. 
The Rademacher complexity of $\F$ w.r.t.\ $\{x_1,\ldots,x_m\} \subset \X$ is 
\[
R_m(\F) =  \frac{1}{m}\E_{\sigma} \sup_{f \in \F}  \sum_{i=1}^{m}\sigma_i f(x_i)
\]
\end{definition}
%
\begin{definition}
Let $\F$ be a family of functions $f:\X\rightarrow\R$ and $\sigma \in \{-1,1\}^m$. 
The Komlos complexity of $\F$ w.r.t.\ $\{x_1,\ldots,x_m\} \subset \X$ is 
\[
K_m(\F) =  \frac{1}{m}\min_{\sigma} \sup_{f \in \F}  \left|\sum_{i=1}^{m}\sigma_i f(x_i)\right|
\]
The Komlos Complexity $K_m(\F)$ without a reference set $\{x_1,\ldots,x_m\}$ is the upper bound on any such set.
\end{definition}

\noindent It well known that $R_m$ is a good measure for generalization when data is chosen uniformly at random (or drawn from an unknown distribution). See \cite{Bartlett:2003:RGC:944919.944944} for other notions of complexity and relationships between them. We claim that $K_m$ is the measure more suitable for active learning, for creating coresets, and for creating mergeable sketches.

@@

%\begin{theorem}
%Let $F(a) = \sum_i a_i$ for $a \in A$ and let $K_m(A) = O(c/m^\gamma)$ for any constant $\gamma \in (0,1]$.
%There exist a coreset for $A$ such that $\tilde{F}(a) = \sum_{i} w_i a_i$, $\operatorname{supp}(w) \le m$ and 
%$$
%\forall \; a \in A \;\;\;  |F(a) - \tilde F(a)| \le \eps n
%$$
%\end{theorem}
%\begin{proof}
%By halving the points repeatedly \el{TODO}.
%\end{proof}
\subsection{General Framework for Coresets}
Consider the function $F = \sum_{i=1}^{m} f(x_i,q)$ and the signed sum of error function $E(q) = \sum_{i=1}^{m} \sigma_i f(x_i,q)$ where $\sigma_i \in \{-1,1\}$.
Now consider $F_1(q) = F(q) + E(q) = \sum_{i=1}^{m} f(x_i,q)  + \sum_{i=1}^{m} \sigma_i f(x_i,q)  = 2 \sum_{i | \sigma_i=1} f(x_i,q)$ and similarly $F_{-1}$. We have that both $F_1(q), F_{-1}(q)$ are approximations for $F(q)$ obtained by coresets of item weights of $2$ and an error at most $|F_1(q) - F(q)| =  |E(q)|$.
%
The above is true for any choice of signs $\sigma_i$, specifically, for those minimizing $\max_q | E(q)|$.
Note that we can select signs such that $|E(q)| \le m K_m(\F)$.
A function is sketch-able if $K_m(\F) = O(m^\gamma)$ for any constant $\gamma \in [0,1)$.
Typically we see that $K_m(\F) = O(c/m)$ or $K_m(\F)= c/\sqrt{m}$ for some $c$ which does not depend on $m$.

\begin{fact}
For any function family $\F$ with a corresponding Komlos Complexity $K_m(\F) = O(c/m)$ there exists a coreset of size 
$O(c/\eps)$ whose error is at most $\eps n$.
\end{fact}
\begin{fact}
For any function family $\F$ with a corresponding Komlos Complexity $K_m(\F) = O(c/\sqrt{m})$ there exists a coreset of size 
$O(c/\eps^2)$ whose error is at most $\eps n$.
\end{fact}

\noindent Proving both facts is trivial. 
At step one, create a coreset of size at most $n/2$ of items of weight $2$ and incur an error of at most $c$ or $c \sqrt{n}$.
Then, we repeat the process and create a coreset of size $n/4$ of items of weight $4$. Here, you incur error of $2c$ or $2c\sqrt{n/2}  = \sqrt{2n}$.
Note that the sum of errors is asymptotically dominated by last iterations. 
Halting the compression at $\Theta(c/\eps)$ or $\Theta(c/\eps^2)$ items respectively achieves the goal.

@@

\subsection{General Framework for Sketching}

@@

\begin{theorem} \label{thm:streaming2}
For any function $f$ with a corresponding Komlos Complexity of $K_m = O(c/\sqrt{m})$ there exists an unbiased fully-mergeable streaming coreset algorithms of size 
$$O\left(c\log^3\left(\min(\eps^2 n/c, \log(n/\delta)) \right)/\eps^2\right)$$ 
whose error is at most $\eps n$ and failure probability at most $\delta$. For $\delta=0$ this results in a deterministic algorithm. The algorithm must be given $\eps, \delta$ as input and does not require a priori knowledge of the stream length $n$.
\end{theorem}
%
%\begin{fact}
%For any function $f$ with a corresponding discrepancy of $\Delta_k = O(\sqrt{k})$ there exists an unbiased fully-mergeable streaming coreset algorithms of size 
%$O(\log^2(\log(1/\delta))/\eps^2)$ whose error is at most $\eps n$ and failure probability at most $\delta$.
%\end{fact}
%
%\begin{fact}
%For any function $f$ with a corresponding discrepancy of $\Delta_k = O(1)$ there exists a fully-mergeable deterministic streaming coreset algorithms of size 
%$O(\log^2(n)/\eps)$ whose error is at most $\eps n$.
%\end{fact}
%\begin{fact}
%For any function $f$ with a corresponding discrepancy of $\Delta_k = O(\sqrt{k})$ there exists a fully-mergeable deterministic streaming coreset algorithms of size 
%$O(\log^2(n)/\eps^2)$ whose error is at most $\eps n$.
%\end{fact}

\section{The Complexity of Analytic Functions of Dot Products}

We provide a coreset suitable for analytical functions of the inner product $\ip{q,x}$ or distance $\|q-x\|$. The idea is to find a set of signs that simultaneously balance $\ip{q,x}^k$ for all powers $k$ and unit vectors $q$. By controlling all powers of $\ip{q,x}$ we control any sum of these powers. It follows that this coreset can be used to control, for example, the logistic loss function $L(q,x) = \log(1+\exp(\ip{q,x}))$ or the gaussian Kernel $K(q,x) = \exp(-\lambda \|q-x\|^2)$.

We start with some notation and trivial properties. 
For a vector $q \in \R^d$ let $q^{\otimes k}$ represent the $k$-dimensional tensor obtained from the outer product of $q$ with itself $k$ times. For a $k$ dimensional tensor with $d^k$ entries $X$ we consider the measure
$\|X\|_{T_k} = \max_{q \in \R^d, \|q\|=1} \left| \langle X, q^{\otimes k}\rangle \right|$.
\begin{fact}
$\|X\|_{T_k}$ is a norm
\end{fact}
\begin{proof}
We prove the claim directly from the definition of a norm.
Notice that for any $X \neq 0$, $\ip{X, q^{\otimes}}$ is a non-zero polynomial in $q$. It follows that there must be $q$ for which its value is non-zero, meaning that $\|X\|_{T_k}=0$ iff $X=0$. For a scalar $a$, we clearly have by definition that
$\|aX\|_{T_k} = |a|\|X\|_{T_k}$.  Lastly, by the max definition we clearly have
$ \|X+Y\|_{T_k} =  \max_q \left| \langle X+Y, q^{\otimes k}\rangle \right| \leq 
\max_q \left| \langle X, q^{\otimes k}\rangle \right| + \max_q\left| \langle Y, q^{\otimes k}\rangle \right| = \|X\|_{T_k} + \|Y\|_{T_k}$
\end{proof}

@@

\begin{lemma}\label{uc}
For any set of vectors $x_i \in \R^d$ there exist a set of signs $\sigma_i$ such that for all $k$ simultaneously $\left\| \sum_i \sigma_i x_i^{\otimes k} \right\|_{T_k} \le O(\sqrt{d k\log^{3}{k}})$ (the $3$ power of the term $\log(k)$ can be reduced to any constant power larger than $2$). 
\end{lemma}
\begin{proof}
The proof will use Banaszczyk's theorem. 
Let $\mathcal K$ be a convex body in Euclidean space with Gaussian measure at least 1/2 ($\Pr[g \in \mathcal K] \ge 1/2$ when $g$ is i.i.d.\ Gaussian).
Let $x_1,\ldots,x_n$ be vectors with $\|x_i\| =O(1)$. 
Then, there exist signs $\sigma_i$ such that $\sum \sigma_i x_i \in C \mathcal K$ for some constant $C$.

@@

A trivial modification of Theorem 1 from \cite{tomioka2014spectral} shows that $\Pr[\|g_k\|_{T_k} \ge c\sqrt{d\log(k)}] \le 1/10k^2$ for some constant $c$. The only change needed in the proof is the size of the epsilon net which changes from $(2\log(3/2)/k)^{kd}$ for \cite{tomioka2014spectral} to $(2\log(3/2)/k)^d$, where the reason we require a net over a smaller space is due to us bounding the inner product with a rank one tensor rather than rank $k$. Union bounding on all values of $k$ we get $\sum_k 1/10k^2 \le 1/2$ which shows $g = [g_1, \operatorname{flat}(g_2), \operatorname{flat}(g_3), \ldots]$ belongs to $\mathcal K$ with probability at least $1/2$. 
%
We now define a mapping $\psi(x)$ of $x\in \R^d$ to a high dimensional space. The function $\operatorname{flat}(X)$ simply strings the entry values of the tensor $X$ into a vector.
$$\psi(x) = [x, \frac{\operatorname{flat}(x^{\otimes 2})}{\sqrt{2\log^2(2)}}, \frac{\operatorname{flat}(x^{\otimes 3})}{\sqrt{3\log^2(3)}}, \ldots,\frac{\operatorname{flat}(x^{\otimes k})}{\sqrt{k\log^2(k)}},\ldots]$$

@@

We are now ready to apply Banaszczyk's theorem. 
There exist signs $\sigma_i$ such that $\psi  = \sum_i \sigma_i \psi(x_i) \in C \mathcal K$.
Since $\psi_k = \sum_i \sigma_i x_i^{\otimes k}/\sqrt{k \log^2{k}}$ we get that $\max_k \|\sum_i \sigma_i  x_i^{\otimes k}\|_{T_k} \le \sqrt{d k \log^{3}(k)}$
\end{proof}

\begin{lemma} \label{lem:komlos anl}
Let $f$ be a function of $\langle x,q\rangle$ such that $f(\langle x,q\rangle) = \sum_k \alpha_k \langle x,q\rangle^k$ is its Taylor expansion. The Komlos Complexity of the Family of functions $f_q(x) = f(\ip{q,x})$, indexed by $\|q\| \leq 1$, is bounded by
\[
K_m = \min_\sigma \sum_i \sigma_i f(x_i,q) =O\left( \sqrt{d} \sum_k  |\alpha_k|\sqrt{k\log^3(k)}\right)
\]
For general $\|q\| \leq R$ we get
\[
K_m = \min_\sigma \sum_i \sigma_i f(x_i,q) =O\left( \sqrt{d} \sum_k  |\alpha_k| R^k \sqrt{ k\log^3(k)}\right)
\]
\end{lemma}
\begin{proof}
The proof follows from combining the above.
$$
\sum_i \sigma_i f(x_i,q) = \sum_k \alpha_k \sum_i \sigma_i \langle x_i,q\rangle^k =  \sum_k \alpha_k  \langle  \sum_i \sigma_i x_i^{\otimes k},q^{\otimes k}\rangle \le \sum_k |\alpha_k| \cdot \| \sum_i \sigma_i x_i^{\otimes k}\|_{T_k} \cdot \|q\|^k
$$
By Lemma \ref{uc} we can find signs $\sigma$ such that 
$\| \sum_i \sigma_i x_i^{\otimes k}\|_{T_k} \le c\sqrt{d k \log^3(k)}$. Substituting into the above, the lemma follows.
\end{proof}

@@

\begin{theorem} \label{thm:analytic2}
Let $f:\R\rightarrow\R$ be analytic. There exist a radius $R$ such that the function family of functions $f_q(x) = f(\|x-q\|^2)$, indexed by $\|q\| \leq \sqrt{R}$, has Komlos complexity of $O(\sqrt{d}/m)$. 
\end{theorem}
\begin{proof}
By transforming $x$ to $\tilde{x} = (1, \sqrt{2}x, \|x\|^2)$ and $q$ to $\tilde{q} = (\|q\|^2, -\sqrt{2}q, 1)$ we get $\ip{\tilde{x},\tilde{q}} = \|q-x\|^2$. Moreover, $\|q\| \le \sqrt{R}$ gives $\|\tilde q\| \le R+1$. The result follows from applying Theorem~\ref{analitic1} to $f(\ip{ \tilde q, \tilde x}) = f(\|q-x\|^2)$.
\end{proof}

\begin{corollary}
The Komlos complexity of the Gaussian kernel $K(q,x) = \exp(-\gamma \|x-q\|^2)$ in dimension $d$ is $O((1+\gamma)\exp(\gamma)\sqrt{d}/m)$.
This improves upon the recent result of \cite{DBLP:journals/corr/abs-1802-01751} by proving the existence of $\eps$ approximation corsets of size $\sqrt{d}/\eps$ for Gaussian kernel density, in the case where $\gamma$ is constant. For the constant $\gamma$ setting this resolves the open problem raised by \cite{DBLP:journals/corr/abs-1802-01751} and matches their lower bound.   
\end{corollary} 
\begin{proof}
W.l.o.g. the maximum distance $\|q-x\|$ is 1. The taylor series of $K$ becomes
$$ \sum_{k=0}^\infty \frac{(-\gamma)^k}{k!} $$
Plugging into the equation in the proof of Theorem~\ref{thm:analytic2} we get that the sum determining the constant is upper bounded by
$$ \sum_{k=0}^\infty \frac{\gamma^{k}\sqrt{ k\log^3(k)}}{k!} = O\left((1+\gamma) \exp(\gamma)\right)$$
\end{proof}
%\end{proof}

@@

From section \ref{} we know that the Komlos complexity of the Gaussian kernel is $K_m = O(\sqrt{d}/m)$. Here, we show that for any positive kernel $K_m = O(1/\sqrt{m})$. This bound is better when $d > m$. More importantly though, there is a very simple, intuitive, and deterministic algorithm for computing the signs $\sigma$. 
Given a collection of data points $X = x_1,\ldots, x_n$ in $\R^d$ the density function $f: \R^d \rightarrow \R$ of a point $q$ is defined as 
$$ f(q) = \sum_{i=1}^{n} K(x_i,q) $$
Here, $K$ is a \emph{positive definite kernel} function, typically based on the distance between $x,y$. The most frequent examples include
$$ K(x,q) = \exp(- \|x-y\|_2^2/\lambda^2)\;\;\; K(x,q) = \exp(- \|x-q\|/\lambda) \; \mbox{and}\;\;\; K(x,y) = (1+\|x-q\|/_2^2/\lambda^2)^{-1}$$
where $\lambda$ is a scaling parameter. For simplicity, we assume that $K(x,x) \leq 1$ for all data points. Notice that for any kernel based on distance we have $K(x,x)=1$ exactly for all $x \in \R^d$.

Using our framework, we need to bound $K_m = \frac{1}{m}\sum_{i=1}^m \sigma_i K(x_i,q)$. For any kernel $K$ there exist a mapping $\phi: \R^d \to {\cal V}$ to an inner product space $\cal V$ such that 
$$ K(x,q) = \ip{\phi(x), \phi(q)} $$
Using this function $\phi$ our objective function becomes
\[
\sum_{i=1}^m \sigma_i K(x_i,q) = \sum_{i=1}^m \sigma_i \ip{\phi(x_i), \phi(q)} =  \ip{ \sum_{i=1}^m \sigma_i \phi(x_i), \phi(q)} \leq  \|\phi(q)\| \cdot \left\|  \sum_{i=1}^m \sigma_i \phi(x_i) \right\| %\leq \left\|  \sum_{i=1}^m \sigma_i \phi(x_i) \right\| 
\]
This upper bound allows us to find values for $\sigma_i$ which bound $K_n$ using a simple algorithm.
Set $\sigma_1 = 1$. Then for $i=2,\ldots,m$ set $\sigma_i = -\operatorname{sign} (\sum_{j=1}^{i-1}\sigma_j  K(x_j, x_i))$.
We will show by induction that $\| \sum_{j=1}^i \sigma_j \phi(x_j) \|^2 \le \sum_{j=1}^i \|\phi(x_j)\|^2$ for all $i$.
This is trivially true for $i=1$. 
Assume by induction that $\| \sum_{j=1}^{i-1} \sigma_j \phi(x_j)\|^2 \le \sum_{j=1}^{i-1} \|\phi(x_j)\|^2$.

>>>>

\begin{definition}
Let $\F$ be a family of functions $f:\X\rightarrow\R$ and $\sigma \in \{-1,1\}^m$ uniformly. 
The Rademacher complexity of $\F$ w.r.t.\ $\{x_1,\ldots,x_m\} \subset \X$ is 
\[
R_m(\F) =  \frac{1}{m}\E_{\sigma} \sup_{f \in \F}  \left| \sum_{i=1}^{m}\sigma_i f(x_i) \right|
\]
\end{definition}
%
In other words, the Radamacher complexity of a family function $\F$ measures how much a function in the family can skew the values when weighted with random signs. The Komlos complexity measures the ability of the family function to skew an adversarial set of signs.

\begin{definition}
Let $\F$ be a family of functions $f:\X\rightarrow\R$ and $\sigma \in \{-1,1\}^m$. 
The Komlos complexity of $\F$ w.r.t.\ $\{x_1,\ldots,x_m\} \subset \X$ is 
\[
K_m(\F) =  \frac{1}{m}\min_{\sigma} \sup_{f \in \F}  \left|\sum_{i=1}^{m}\sigma_i f(x_i)\right|
\]
The Komlos Complexity $K_m(\F)$ without a reference set $\{x_1,\ldots,x_m\}$ is the upper bound on any such set.
\end{definition}

\noindent It well known that $R_m$ is a good measure for generalization when data is chosen uniformly at random (or drawn from an unknown distribution). See \cite{Bartlett:2003:RGC:944919.944944} for other notions of complexity and relationships between them. We claim that $K_m$ is the measure more suitable for \zk{I think we should remove `active learning' from the list} creating coresets, and for creating mergeable sketches.

@@

 \zk{I actually am still having problems with justifying the name Komlos complexity. It might as well be called Banaszczyk's complexity or simply discrepancy complexity. In particular if you could prove Komlos' conjecture it wouldn't affect any of our results, nor simplify our proofs.. Either way, below is my suggested motivation for the definition}
To understand the motivation of the the Komlos Complexity, consider the following informal explanation of Radamacher Complexity applied to ML problems. In this ML problem there exists a distribution over $(x,y)$ pairs where we aim to predict $y$ given $x$. Our goal is to find a predictor from a family of prediction functions that suffers the least loss. A predictor maps to a loss value on each $x,y$ pair so we can discuss the problem of finding a function $f$ from a family $\F$ minimizing $\E_{x,y} f(x,y)$ where $f$ is the loss associated with the corresponding predictor on the $x,y$ pairs. Having a low Radamacher complexity means that if we sample $m$ pairs, split them into two uniformly at random and find the function $f$ minimizing the loss over the first half, the loss over the second half is bounded. This translates to a generalization bound stating that the function $f$ minimizing the loss over a uniform random sample of $(x,y)$ pairs, will perform well on the entire distribution. In other words, the Radamacher complexity gives a guarantee for the distribution loss of the minimizer of a finite set of pairs chosen uniformly at random.

Coming back to Komlos Complexity, having the ability to choose the signs in a non-uniform way lets us choose a subset of pairs $(x,y)$ such that the minimizer $f$ over the subset is guaranteed to have (roughly) the same performance on the entire set. This allows us to obtain a guarantee for the distribution loss of a minimizer of a finite set of pairs chosen in an adaptive way to the family function $\F$. This set is in fact a core-set and the Komlos complexity helps us determine that possible size of a coreset. The tradeoff of considering a coreset vs.\ a random sample is that on one hand, it is clearly much simpler to obtain a random sample. On the other hand we will show in the following sections several examples for which a coreset can be significantly smaller than the random sample while maintaining the same guarantees. This will be done by showing that for a wide range of interesting problems in machine learning $K_m(\F) = o(R_m(\F))$.

%\begin{theorem}
%Let $F(a) = \sum_i a_i$ for $a \in A$ and let $K_m(A) = O(c/m^\gamma)$ for any constant $\gamma \in (0,1]$.
%There exist a coreset for $A$ such that $\tilde{F}(a) = \sum_{i} w_i a_i$, $\operatorname{supp}(w) \le m$ and 
%$$
%\forall \; a \in A \;\;\;  |F(a) - \tilde F(a)| \le \eps n
%$$
%\end{theorem}
%\begin{proof}
%By halving the points repeatedly \el{TODO}.
%\end{proof}
\subsection{General Framework for Coresets}
In what follows it will be convenient to represent the function family as a family of vectors representing the functions. The vector family will be denoted as $Q$, where we use $q$ to denote a single vector corresponding to a function. We a universal function $f$ such that $f(x,q)$ corresponds to the value of the function matching $q$ when applied to $x$. We will assume from hereon that for any $x,q$ $|f(x,q)| \leq 1$; this allows to simplify the presentation. Our results can extend to $|f(x,q)| \leq R$ and will have dependencies on the bound $R$. Achieving these bounds is a purely technical task that we defer for the sake of readability.

Consider the function $F(q) = \sum_{i=1}^{m} f(x_i,q)$ and the signed sum of error function $E(q) = \sum_{i=1}^{m} \sigma_i f(x_i,q)$ where $\sigma_i \in \{-1,1\}$.
Now consider $F_1(q) = F(q) + E(q) = \sum_{i=1}^{m} f(x_i,q)  + \sum_{i=1}^{m} \sigma_i f(x_i,q)  = 2 \sum_{i | \sigma_i=1} f(x_i,q)$ and similarly $F_{-1}(q)$. We have that both $F_1(q), F_{-1}(q)$ are approximations for $F(q)$ obtained by coresets of item weights of $2$ and an error at most $|F_1(q) - F(q)| =  |E(q)|$.
%
The above is true for any choice of signs $\sigma_i$, specifically, for those minimizing $\max_q | E(q)|$.
Note that we can select signs such that $|E(q)| \le m K_m(\F)$.
A function is sketch-able if $K_m(\F) = o(m)$.
Typically we see that $K_m(\F) = O(c/m)$ or $K_m(\F)= c/\sqrt{m}$ for some $c$ which does not depend on $m$.

\begin{fact}
For any function family $\F$ with a corresponding Komlos Complexity $K_m(\F) = O(c/m)$ there exists a coreset of size 
$O(c/\eps)$ whose error is at most $\eps n$.
\end{fact}
\begin{fact}
For any function family $\F$ with a corresponding Komlos Complexity $K_m(\F) = O(c/\sqrt{m})$ there exists a coreset of size 
$O(c^2/\eps^2)$ whose error is at most $\eps n$.
\end{fact}

\noindent Proving both facts is trivial. 
At step one, create a coreset of size at most $n/2$ of items of weight $2$ and incur an error of at most $c$ or $c \sqrt{n}$.
Then, we repeat the process and create a coreset of size $n/4$ of items of weight $4$. Here, you incur error of $2c$ or $2c\sqrt{n/2}  = c\sqrt{2n}$.
Note that the sum of errors is a geometric sequence that is asymptotically dominated by its last element. 
Halting the compression at $\Theta(c/\eps)$ or $\Theta(c^2/\eps^2)$ items respectively achieves the goal.

@@

\subsection{General Framework for Sketching} \label{sec:sketch}

@@

\begin{theorem} \label{thm:streaming2}
For any function $f$ with a corresponding Komlos Complexity of $K_m = O(c/\sqrt{m})$ there exists an unbiased fully-mergeable streaming coreset algorithms of size 
$$O\left(c^2\log^3\left(\min(\eps^2 n/c, \log(n/\delta)) \right)/\eps^2\right)$$ 
whose error is at most $\eps n$ and failure probability at most $\delta$. For $\delta=0$ this results in a deterministic algorithm. The algorithm must be given $\eps, \delta$ as input and does not require a priori knowledge of the stream length $n$.
\end{theorem}
%
%\begin{fact}
%For any function $f$ with a corresponding discrepancy of $\Delta_k = O(\sqrt{k})$ there exists an unbiased fully-mergeable streaming coreset algorithms of size 
%$O(\log^2(\log(1/\delta))/\eps^2)$ whose error is at most $\eps n$ and failure probability at most $\delta$.
%\end{fact}
%
%\begin{fact}
%For any function $f$ with a corresponding discrepancy of $\Delta_k = O(1)$ there exists a fully-mergeable deterministic streaming coreset algorithms of size 
%$O(\log^2(n)/\eps)$ whose error is at most $\eps n$.
%\end{fact}
%\begin{fact}
%For any function $f$ with a corresponding discrepancy of $\Delta_k = O(\sqrt{k})$ there exists a fully-mergeable deterministic streaming coreset algorithms of size 
%$O(\log^2(n)/\eps^2)$ whose error is at most $\eps n$.
%\end{fact}

\section{The Complexity of Analytic Functions of Dot Products} \label{sec:analytic}

Now that we proved the usefulness of the Komlos Complexity we move to upper bound it for interesting family functions. We provide a coreset suitable for analytical functions of the inner product $\ip{q,x}$ or Euclidean distance $\|q-x\|$.

The idea is to find a set of signs that simultaneously balance $\ip{q,x}^k$ for all powers $k$ and unit vectors\footnote{We Assume that $\|x\|,\|q\| \leq 1$ for ease of presentation. As above our results extend to generic bounds on the radius of $q$} $q$. By controlling all powers of $\ip{q,x}$ we control any sum of these powers. It follows that this coreset can be used to control, for example, the logistic loss function $L(q,x) = \log(1+\exp(\ip{q,x}))$ or the gaussian Kernel $K(q,x) = \exp(-\lambda \|q-x\|^2)$.

We start with some notation and trivial properties. 
For a vector $q \in \R^d$ let $q^{\otimes k}$ represent the $k$-dimensional tensor obtained from the outer product of $q$ with itself $k$ times. For a $k$ dimensional tensor with $d^k$ entries $X$ we consider the measure
$\|X\|_{T_k} = \max_{q \in \R^d, \|q\|=1} \left| \langle X, q^{\otimes k}\rangle \right|$.
\begin{fact}
$\|X\|_{T_k}$ is a norm
\end{fact}
\begin{proof}
We prove the claim directly from the definition of a norm.
Notice that for any $X \neq 0$, $\ip{X, q^{\otimes}}$ is a non-zero polynomial in $q$. It follows that there must be $q$ for which its value is non-zero, meaning that $\|X\|_{T_k}=0$ iff $X=0$. For a scalar $a$, we clearly have by definition that
$\|aX\|_{T_k} = |a|\|X\|_{T_k}$.  Lastly, by the max definition we  have
$ \|X+Y\|_{T_k} =  \max_q \left| \langle X+Y, q^{\otimes k}\rangle \right| \leq 
\max_q \left| \langle X, q^{\otimes k}\rangle \right| + \max_q\left| \langle Y, q^{\otimes k}\rangle \right| = \|X\|_{T_k} + \|Y\|_{T_k}$
\end{proof}

@@

\begin{lemma}\label{uc}
For any set of vectors $x_i \in \R^d$ with $\|x_i\| \leq 1$ there exist a set of signs $\sigma_i$ such that for all $k$ simultaneously $\left\| \sum_i \sigma_i x_i^{\otimes k} \right\|_{T_k} \le O(\sqrt{d k\log^{3}{k}})$ (the $3$ power of the term $\log(k)$ can be reduced to any constant power larger than $2$). 
\end{lemma}
\begin{proof}
The proof will use Banaszczyk's theorem \cite{Banaszczyk}. 
Let $\mathcal K$ be a convex body in Euclidean space with Gaussian measure at least 1/2 ($\Pr[g \in \mathcal K] \ge 1/2$ when $g$ is i.i.d.\ Gaussian).
Let $x_1,\ldots,x_n$ be vectors with $\|x_i\| \leq 1$. 
Then, there exist signs $\sigma_i$ such that $\sum \sigma_i x_i \in C \mathcal K$ for some constant $C$.

@@

A trivial modification of Theorem 1 from \cite{tomioka2014spectral} shows that $\Pr[\|g_k\|_{T_k} \ge c\sqrt{d\log(k)}] \le 1/10k^2$ for some constant $c$. The only change needed in the proof is the size of the epsilon net which changes from $(2\log(3/2)/k)^{kd}$ for \cite{tomioka2014spectral} to $(2\log(3/2)/k)^d$, where the reason we require a net over a smaller space is due to us bounding the inner product with a rank one tensor rather than rank $k$. Union bounding on all values of $k$ we get $\sum_k 1/10k^2 \le 1/2$ which shows $g = [g_1, \operatorname{flat}(g_2), \operatorname{flat}(g_3), \ldots]$ belongs to $\mathcal K$ with probability at least $1/2$, where $\operatorname{flat}(g_k)$ is the flattening of the tensor into a one dimensional vector. 
%
We now define a mapping $\psi(x)$ of $x\in \R^d$ to a high dimensional space. The function $\operatorname{flat}(X)$ simply strings the entry values of the tensor $X$ into a vector.
$$\psi(x) = \left[x, \frac{\operatorname{flat}(x^{\otimes 2})}{\sqrt{2\log^2(2)}}, \frac{\operatorname{flat}(x^{\otimes 3})}{\sqrt{3\log^2(3)}}, \ldots,\frac{\operatorname{flat}(x^{\otimes k})}{\sqrt{k\log^2(k)}},\ldots \right]$$

@@

We are now ready to apply Banaszczyk's theorem. 
There exist signs $\sigma_i$ such that $\psi  = \sum_i \sigma_i \psi(x_i) \in C \mathcal K$.
Since $\psi_k = \sum_i \sigma_i x_i^{\otimes k}/\sqrt{k \log^2{k}}$ we get that 
$$\max_k \frac{\|\sum_i \sigma_i  x_i^{\otimes k}\|_{T_k}}{\sqrt{k \log^{3}(k)}} \le O\left( \sqrt{d} \right)$$
\end{proof}

\begin{lemma} \label{lem:komlos anl}
Let $f$ be a function of $\langle x,q\rangle$ such that $f(\langle x,q\rangle) = \sum_k \alpha_k \langle x,q\rangle^k$ is its Taylor expansion. The Komlos Complexity of the Family of functions $f_q(x) = f(\ip{q,x})$, indexed by $\|q\| \leq 1$, is bounded by
\[
K_m = \min_\sigma \sum_i \sigma_i f(x_i,q) =O\left( \sqrt{d} \sum_k  |\alpha_k|\sqrt{k\log^3(k)}\right)
\]
For general $\|q\| \leq R$ we get
\[
K_m = \min_\sigma \sum_i \sigma_i f(x_i,q) =O\left( \sqrt{d} \sum_k  |\alpha_k| R^k \sqrt{ k\log^3(k)}\right)
\]
\end{lemma}
\begin{proof}
The proof follows from combining the above.
$$
\sum_i \sigma_i f(x_i,q) = \sum_k \alpha_k \sum_i \sigma_i \ip{ x_i,q}^k =  \sum_k \alpha_k  \ip{  \sum_i \sigma_i x_i^{\otimes k},q^{\otimes k}} \le $$
$$\sum_k |\alpha_k| \cdot \left\| \sum_i \sigma_i x_i^{\otimes k}\right\|_{T_k} \cdot \|q\|^k
$$
By Lemma \ref{uc} we can find signs $\sigma$ such that 
$\left\| \sum_i \sigma_i x_i^{\otimes k}\right\|_{T_k} \le c\sqrt{d k \log^3(k)}$. Substituting into the above, the lemma follows.
\end{proof}

@@

\begin{theorem} \label{thm:analytic2}
Let $f:\R\rightarrow\R$ be analytic. There exist a radius $R$ such that the function family of functions $f_q(x) = f(\|x-q\|^2)$, indexed by $\|q\| \leq R$, has Komlos complexity of $O(\sqrt{d}/m)$. 
\end{theorem}
\begin{proof}
By transforming $x$ to $\tilde{x} = (1, \sqrt{2}x, \|x\|^2)$ and $q$ to $\tilde{q} = (\|q\|^2, -\sqrt{2}q, 1)$ we get $\ip{\tilde{x},\tilde{q}} = \|q-x\|^2$. Moreover, $\|q\| \le R$ gives $\|\tilde q\| \le R^2+1$. The result follows from applying Theorem~\ref{analitic1} to $f(\ip{ \tilde q, \tilde x}) = f(\|q-x\|^2)$.
\end{proof}

\begin{corollary}
The Komlos complexity of the Gaussian kernel $K(q,x) = \exp(-\gamma \|x-q\|^2)$ in dimension $d$ is $O((1+\gamma)\exp(\gamma)\sqrt{d}/m)$.
\end{corollary} 
This improves upon the recent result of \cite{DBLP:journals/corr/abs-1802-01751} by proving the existence of $\eps$ approximation corsets of size $\sqrt{d}/\eps$ for Gaussian kernel density, in the case where $\gamma$ is constant. For the constant $\gamma$ setting this resolves the open problem raised by \cite{DBLP:journals/corr/abs-1802-01751} and matches their lower bound.

\begin{proof}
W.l.o.g. the maximum distance $\|q-x\|$ is 1. The taylor series of $K$ becomes
$$ \sum_{k=0}^\infty \frac{(-\gamma)^k}{k!} $$
Plugging into the equation in the proof of Theorem~\ref{thm:analytic2} we get that the sum determining the constant is upper bounded by
$$ \sum_{k=0}^\infty \frac{\gamma^{k}\sqrt{ k\log^3(k)}}{k!} = O\left((1+\gamma) \exp(\gamma)\right)$$
\end{proof}
%\end{proof}

@@

From section \ref{sec:analytic} we know that the Komlos complexity of the Gaussian kernel is $K_m = O(\sqrt{d}/m)$. The result however is non-constructive in the sense that we are not aware of an algorithm that can obtain the $\sigma$ signs efficiently. 
Here, we provide a computationally efficient bound that can be achieved with a straightforward algorithm of complexity $O(m^2)$. Together with the results of Section~\ref{sec:sketch} this provides an efficient sketching algorithm for Kernel Density Estimation. We show that for any positive kernel $K_m = O(1/\sqrt{m})$. This bound is superior to that of the previous section when $d > \sqrt{m}$. More importantly though, there is a very simple, intuitive, and deterministic algorithm for computing the signs $\sigma$. 
Given a collection of data points $X = x_1,\ldots, x_n$ in $\R^d$ the density function $f: \R^d \rightarrow \R$ of a point $q$ is defined as 
$$ f(q) = \sum_{i=1}^{n} K(x_i,q) $$
Here, $K$ is a \emph{positive definite kernel} function, typically based on the distance between $x,y$. The most frequent examples include
$$ K(x,q) = \exp(- \|x-y\|_2^2/\lambda^2)\;\;\; K(x,q) = \exp(- \|x-q\|/\lambda) \; \mbox{and}\;\;\; K(x,y) = (1+\|x-q\|/_2^2/\lambda^2)^{-1}$$
where $\lambda$ is a scaling parameter. For simplicity, we assume that $K(x,x) \leq 1$ for all data points. Notice that for any kernel based on distance we have $K(x,x)=1$ exactly for all $x \in \R^d$.

The algorithm we proposed works as follows: Set $\sigma_1 = 1$. For $i=2,\ldots,m$ set $\sigma_i = -\operatorname{sign} (\sum_{j=1}^{i-1}\sigma_j  K(x_j, x_i))$.
\begin{theorem}
The algorithm above achieves a set of signs $\sigma_1,\ldots,\sigma_m$ such that for any $q$,
$$ \frac{1}{m}\sum_{i=1}^m \sigma_i K(x_i,q) \leq \sqrt{m} $$
\end{theorem}
